Risk Perceptions for Wearable Devices

### 2015-03-04 tex-inputs/intro.tex
finished intro and did a bit of tweaks to related work

Wearable technologies, or ``wearables,'' are a \$700 million industry \cite{cmo} consisting of electronically enhanced clothing items and accessories aims to interweave technological interaction into everyday life. A top 25 market research company estimates that 52\% of technology consumers are aware of wearables and 33\% said they were likely to buy one~\cite{NPD}. With 20\% of the general population owning at least one wearable and 10\% using it daily~\cite{WearableStatNews}, wearables are transforming ubiquitous computing into a part of every day life. Forbes has named 2014 the ``Year of Wearable Technology \cite{Forbes}.''

@@

Along with these benefits, wearable devices bring with them new potential privacy issues which expose users' activities without their awareness or consent. Fitbit allowed sex to be tracked as exercise while fitness profiles were public by default~\cite{Fitbit}, resulting in the inadvertent disclosure of sensitive information. Public discomfort toward facial recognition was so strong that the capability was not even deployed for Google Glass \cite{GlassDetection}. Google Glass, an iconic wearable of today, has since disappeared \cite{13_google_2015}. Most suspect that the reason for its disappearance was because it was the ``straw the broke the back of the privacy camel's back'' \cite{14_dvorak_2014}. Some Glass wearers faced assault \cite{1_russell_2014, 15_mashable_2014, 16_gross_2014}, while bystanders felt uncomfortable personal details, conversations, and photos possibly recorded.

{\color {red} smart phone concerns. we haven't gotten it right for this yet! We need to do something for wearables before it's too late.}

{\color {red} We should warn people about such things. But we can't do that, since attention is a finite resource. This calls for privacy research in this area!}  To avoid scandalous breaches of privacy and public opposition to new capabilities, it is critical we understand user concerns surrounding wearable devices to prevent negative social impact and breaches of privacy. There has been extensive work on the need for showing users and bystanders how certain data is being captured so that they can have more control. We have an issue of showing everyone everything all the time, since attention is a finite resource. We need to identify situations that people want to warn them about, and transparently access things that are benign. A better understanding of users' risk perceptions will enable researchers and companies to focus on users' concerns. The goal of this paper is to gain a better sense of what those concerns are.

{\color {red} better intro to here} We surveyed 1,784 Internet users for their perceptions of wearable devices. In this work, we contribute the following: \\[-0.8cm]

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item We compare users' perceptions of a range of privacy and security risks of wearables. Users care much more about the type of data, than the recipient of the data.
\item We observed that users make little distinction between sharing data with friends, co-workers, or the general public, but are relatively comfortable with an application's servers receiving their data.
\item Our participants viewed data-collection capabilities of wearable devices as benign compared to more familiar technologies. However, we suspect that this may be due to a lack of exposure to these newer technologies.
%\item We report people's self-reported top concerns for wearable devices. Privacy, by far, is the top concern. Other notable risks include information security, long-term health risks from use, high financial cost, and change in social norms. 
\end{itemize}

>>>>

Wearable technologies, or ``wearables,'' are a \$700 million industry \cite{cmo} of electronically enhanced clothing items and accessories that interweaves technological interaction with everyday life. A top 25 market research company estimates that 52\% of technology consumers are aware of wearables and 33\% said they were likely to buy one~\cite{NPD}. With 20\% of the general population owning at least one wearable and 10\% using it daily~\cite{WearableStatNews}, wearables are transforming ubiquitous computing into a part of every day life. Forbes has named 2014 the ``Year of Wearable Technology \cite{Forbes}.''

@@

Along with these benefits, wearable devices bring with them new potential privacy and security risks which expose users' activities without their awareness or consent. Fitbit allowed sex to be tracked as exercise while fitness profiles were public by default~\cite{Fitbit}, resulting in the inadvertent disclosure of sensitive information. Public discomfort toward facial recognition prevented the capability from being deployed to Google Glass \cite{GlassDetection}. Google Glass, the iconic wearable of its time, has since disappeared \cite{13_google_2015}. Most suspect that the reason for its disappearance was because it was the ``straw the broke the back of the privacy camel's back'' \cite{14_dvorak_2014}. Some Glass wearers faced assault \cite{1_russell_2014, 15_mashable_2014, 16_gross_2014}, while bystanders felt uncomfortable personal details, conversations, and photos possibly recorded.

We have seen similar privacy \cite{kelley2013privacy, sadeh2009understanding, shklovski2014leakiness} and security issues \cite{enck2011study, felt2011survey} related to data capture with respect to smartphones. Mobile platforms have tried to address this by communicating data capture to users as the data is captured. However, many users are habituated to these notifications, because they see them all the time, often for things that they don't care about \cite{felt2012android}.

Human attention is a finite resource \cite{bohme2011security}. Given the additional capabilities of wearable devices and their increasing popularity, we need to study user concerns to warn users only about situations they are likely to care about. Informed, select notifications make for a better user experience and prevents habituation to such notifications while avoid scandalous breaches of privacy.

The goal of this paper is to gain a better sense of user concerns for wearables. To our knowledge, this is the first large-scale study to investigate user security and privacy concerns for wearable devices. We surveyed 1,784 Internet users for their perceptions of wearable devices and contribute the following: \\[-0.8cm]

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item Comparisons of users' perceptions of a range of privacy and security risks of wearables. We found that users care much more about the type of data than the recipient of the data.
\item Insight into how users feel about various data recipients. We observed that users make less distinction between sharing data with friends, co-workers, and the general public, comparatively to sharing with an application's servers.
\item Rankings of how data-collection capabilities of wearable devices compared to more familiar technologies.  Most saw new capabilities as benign, but we suspect that this may be due to a lack of exposure to these newer technologies.
\item A taxonomy and report of users' self-reported top concerns for wearable devices. Privacy is by far the top concern, along with security, health risks, financial risk, and social impact. 
\end{itemize}

### 2015-03-04 tex-inputs/related_work.tex
finished intro and did a bit of tweaks to related work

\subsection{Ubiquitous Sensing}
Many authors have emphasized that we are rapidly moving towards a world of ubiquitous sensing and data capture, with ensuing privacy challenges~\cite{abowd2000charting,palen2003unpacking,camp2000internet}. Many researchers have worked to study how privacy can be preserved in such a future. Examples of such efforts include frameworks to design for privacy~\cite{bellotti1993design,camp2003designing,langheinrich2001privacy} or evaluate privacy~\cite{scholtz2004toward} in ubiquitous computing applications. Others have suggested various models for understanding privacy in ubiquitous computing systems~\cite{hong2004privacy, jiang2002approximate}.
However, none of these works attempted to quantify or rank user concern over different privacy risks.

\subsection{Smartphones and Wearables Concerns}
Many researchers have attempted to study end-user concerns about security or privacy issues associated with their smartphones~\cite{chin2012measuring, palen2000going, felt2012android}.

Recently Denning et. al studied privacy concerns bystanders have when others around them are wearing a wearable device~\cite{Denning2014}. Participants in their study expressed concerns due to the nature of wearables (subtle UI and potenitally ubiquitous). While their research examined the privacy concerns of bystanders in the presence of wearables, we are aware of no work that has looked at the privacy concerns of owners.

\subsection{User Perceptions and Behaviors}
In this paper we focus on measuring people's perceptions of security and privacy risks.
One limitation of user perceptions is that people don't always have enough information to make privacy-sensitive decisions; even if they do, they often trade off long-term privacy for short-term benefits~\cite{acquisti2005privacy}. Also, actual behavior may deviate from self-reported behaviors~\cite{jensen2005privacy} and privacy preferences~\cite{spiekermann2001privacy}.

>>>>

\subsection{Wearables Concerns}
A small-scale interview of how bystanders feel about wearable devices \cite{denning2014situ} found that bystanders were predominantly split between having indifferent and negative reactions to the device. A variety of factors that make recording more or less acceptable, including what they are doing when the recording is being taken.  Additionally, bystanders are expressed interest in being able to give permissions for the data being captured. We also investigate how the type and mechanism of data capture affects privacy concerns, but we examine the privacy concerns of wearables owners at a large-scale while their research examined the privacy concerns of wearables bystanders at a small-scale.

\subsection{Ubiquitous Sensing Concerns}
Many authors have emphasized that we are rapidly moving towards a world of ubiquitous sensing and data capture, with ensuing privacy challenges~\cite{abowd2000charting,palen2003unpacking,camp2000internet}. Many researchers have worked to study how privacy can be preserved in such a future. Examples of such efforts include frameworks to design for privacy~\cite{bellotti1993design,camp2003designing,langheinrich2001privacy} or evaluate privacy~\cite{scholtz2004toward} in ubiquitous computing applications. Others have suggested various models for understanding privacy in ubiquitous computing systems~\cite{hong2004privacy, jiang2002approximate}. However, none of these works attempted to quantify or rank user concern over different privacy risks. 
%one-sentence on takeaways from these sources and comment on them!

\subsection{Smartphones Concerns}
Many researchers have attempted to study end-user concerns about security or privacy issues associated with their smartphones~\cite{chin2012measuring, palen2000going, felt2012android}.  
%one-sentence on takeaways from these sources and comment on them!

\subsection{User Perceptions and Behaviors}
In this paper we focus on measuring people's perceptions of security and privacy risks.
One limitation of user perceptions is that people don't always have enough information to make privacy-sensitive decisions; even if they do, they often trade off long-term privacy for short-term benefits~\cite{acquisti2005privacy}. Also, actual behavior may deviate from self-reported behaviors~\cite{jensen2005privacy} and privacy preferences~\cite{spiekermann2001privacy}. 
%one-sentence on takeaways from these sources and comment on them!
